An Efficient Minibatch Acceptance Test for MCMC

### 2016-07-06 nips_2016.tex
updates to first section

In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$, we
can write
\begin{equation}\label{eq:relationship}
\Delta^* = \Delta + X_{\rm norm}, \quad X_{\rm norm} \sim \mathcal{\bar{N}}(0, \sigma^2(\Delta^*)),
\end{equation}
where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is approximately
normal with variance $\sigma^2(\Delta^*)$. 
But to perform the test in equation (\ref{theexacttest}) we want $\Delta + X$ for
a logistic random $X$ (call it $X_{\rm logistic}$ from now on). 
Our approach is to decompose $X_{\rm logistic}$ as
\begin{equation}\label{eq:deconvolution}
X_{\rm logistic} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm}$ is a zero-mean Gaussian and $X_{\rm corr}$ is a zero-mean
``correction'' term.  These two add (i.e., their distributions convolve) to form $X_{\rm logistic}$. The acceptance
test is now:
\begin{equation}\label{eq:criteria}
\Delta + X_{\rm logistic} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0, \sigma^2(\Delta^*))$,
the above test is exact as well. And as we will see later,  if there is a maximum error $\epsilon$ between
the CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0, \sigma^2(\Delta^*))$,
then the acceptance test has an error of at most $\epsilon$.

>>>>

In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$, we
can write
\begin{equation}\label{eq:relationship}
\Delta^* = \Delta + X_{\rm norm}, \quad X_{\rm norm} \sim \mathcal{\bar{N}}(0, \sigma^2(\Delta^*)),
\end{equation}
where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is approximately
normal with variance $\sigma^2(\Delta^*)$. 
But to perform the test in equation (\ref{theexacttest}) we want $\Delta + X$ for
a logistic random $X$ (call it $X_{\rm logistic}$ from now on). 
Our approach is to decompose $X_{\rm logistic}$ as
\begin{equation}\label{eq:deconvolution}
X_{\rm logistic} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm}$ is a zero-mean Gaussian and $X_{\rm corr}$ is a zero-mean
``correction'' term.  These two add (i.e., their distributions convolve) to form $X_{\rm logistic}$. The acceptance
test is now:
\begin{equation}\label{eq:criteria}
\Delta + X_{\rm logistic} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
i.e. given an approximate estimate $\Delta^*$ from the current data minibatch (assuming its
variance is small enough) we test acceptance by adding a random variable $X_{\rm corr}$ and
then accept $\theta'$ if the result is positive.

If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0, \sigma^2(\Delta^*))$,
the above test is exact as well. And as we will see later,  if there is a maximum error $\epsilon$ between
the CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0, \sigma^2(\Delta^*))$,
then the acceptance test has an error of at most $\epsilon$.

\begin{figure}[t]
    \centering
    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
    \caption{
    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
    standard logistic distribution. We use three standard deviation values of $X_{\rm norm}$. The
    two red curves convolve to form the logistic, etc. The $y$-axis is capped at 0.5 for
    readability. This figure must be viewed in color.
    }
    \label{fig:deconvolution}
    %\vspace{-10pt}
\end{figure}
